Layer dependence as a measure of local dependence

### 2015-04-21 Response to referees.tex

\documentclass[a4paper, 12pt]{report}
\setlength\arraycolsep{2pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}
\setcounter{tocdepth}{0}
\usepackage{graphicx}
\usepackage{chicago}
\usepackage{amsfonts}
\usepackage{pdfpages}
\bibliographystyle{chicago}

\newcommand{\eref}[1]{(\ref{#1})}
\newcommand{\fref}[1]{Figure \ref{#1}}
\newcommand{\sref}[1]{\S\ref{#1}}
\newcommand{\tref}[1]{Table \ref{#1}}
\newcommand{\aref}[1]{\ref{#1}}
\newcommand{\bi}{\begin{itemize}}
\renewcommand{\i}{\item}
\newcommand{\ei}{\end{itemize}}

\begin{document}

\section*{Response to Referee A comments}

Referee A provided a number of helpful, specific comments on how to improve the paper and the ideas behind layer dependence. We have made significant revisions to our paper to address Referee A's comments. Some of our revisions were not specifically brought up by Referee A but were inspired by Referee A's comments. For example we have formalised and discussed in detail several of the concepts and properties of layer dependence in the paper, e.g. section 2.
Our responses to specific comments from Referee A are:

\begin{enumerate}

\item We agree with the potential confusion caused by $(u>a)$ and we have replaced it with the more common indicator function $I_\alpha(u)$.

As we are dealing mostly with random variables in the paper, the use of lower case is not likely to cause confusion with realisations of the random variables. In addition we have made it clear if the variables are random as opposed to constants.

\item	We agree and we have named some of the copulas illustrated in Figure 1. Remaining copulas have been specially constructed to exhibit specific dependence structures, i.e. not from any well-known family. Our view is that the main aim of Figure 1 is to illustrate how layer dependence can capture a range of dependence structures, and the actual form of the copula used is of secondary interest.

\item	We agree and we have included Table 1 to illustrate how layer dependence changes for various copulas (independence, perfect dependence) and their transformations.

\item	We agree and we have replaced the expectations with probabilities. We have also inserted the suggested reference.

\item	We agree and we have used indicator functions to express the tail concentration function. We have also included the suggested reference.

\item This was a typographical error and we have removed the comment.

\item	We agree and we have removed the expression.

\item	We agree that many copulas can satisfy the same fitted layer dependence curve. We realise that copula fitting based on layer dependence involves a detailed discussion and warrants a separate paper in order to be properly covered. Hence we have decided to remove this discussion from the paper and instead focus on a complete discussion of layer dependence and its propertes.

\item	We agree and we have included more detailed discussion in section 9.

\item	Please refer to comment 8. We have removed the discussion on copula fitting based on layer dependence from this paper.

\item	Please refer to comment 8. We have removed the discussion on copula fitting based on layer dependence from this paper.

\item	We agree and we have amended the wording in the section.

\end{enumerate}

\newpage

\section*{Response to Referee B comments}

We can see Referee B's point of view that the proposed layer dependence may not have been set out in a logical sequence or explained in sufficient detail. Hence we have included a significant amount of preliminary detail and discussion which leads on to the definition of layer dependence. Hopefully this will explain the motivation and usefulness of layer dependence.

Our responses to Referee B's specific comments are:

\begin{enumerate}

\item	Regarding the first comment, we have included additional explanation of how layer dependence captures the dependence structure for each copula illustrated in Figure 1, i.e. by reflecting the scatter of points continuously along the 45 degree line. We are in a way trying to ``catch tail and middle simultaneously", and we are doing so for an infinite number of combinations as alpha varies from 0 to 1.

\item	Regarding the second comment, we understand that significant discussion is required for the factor copula model and how it can be used to model any layer dependence curve. Hence we have removed this discussion so that the current paper focuses on explaining layer dependence and its use as a local dependence measure. The application of layer dependence to copula fitting is best discussed in a separate paper.

\item	Regarding the third comment, we agree and we have used the more common notation $I_\alpha(u)$ for indicator variables.

\end{enumerate}

\end{document}

### 2015-04-21 ld10.tex
more changes

Dependence between two random variables generally varies with percentile. For example extreme movements in two stock markets are likely to be highly related  whereas minor fluctuations may be relatively independent. Natural catastrophes create significant insurance losses for several classes of business at the same time, while attritional losses between various classes are weakly dependent.

@@

This paper introduces, illustrates and analyzes an alternate local dependence measure called  ``layer dependence." Layer dependence is the covariance between a random variable and a single ``layer" of another. Layer dependence is also the ``gap" between upper and lower conditional tail expectations. Random variables are replaced with their percentile rank transforms and layer dependence is calculated entirely from the copula underlying the joint distribution. Hence of interest is rank dependence rather than dependence between random variables in their original scale, as the latter is often distorted by marginal distributions.

@@

Therefore calculating layer dependence at the first instance extracts essential and interpretable information -- the dependence structure. For a parametric copula, the implication of its parametric form and parameters on the dependence structure is not always apparent. Similarly computing layer dependence from data guides the selection and fitting of an appropriate copula.

@@

This paper combines the concepts of rank correlation and loss layers to form a local dependence measure called layer dependence. The following explains these ideas and the definition of layer dependence sequentially.

A familiar construct in the study of bivariate dependence is Spearman's correlation \citep{embrechts2002correlation}, the linear correlation between ranks of two random variables. Rank dependence avoids distortion from marginal distributions unlike with for example Pearson's correlation \citep{mcneil2005qrm} which measures the degree of linear relationship between random variables in their original values. Spearman's correlation can also be applied to estimate copula parameters using the method of moments (\cite{kojadinovic2010comparison}, \cite{bouye2000copulas}). However Spearman's correlation suffers from shortcomings and, as an aggregate measure, may be inappropriate for assessing local dependence which varies across the joint distribution including the tails.

Another familiar construct, in reinsurance, is a loss layer \citep{wang1995insurance}.  For example the 95\%-96\% layer of a random continuous loss $x$ is the portion of $x$ between its 95th and 96th percentile
$$
\min\{(x-x_{0.95})^+,x_{0.96}-x_{0.95}\}\ ,
$$
where $()^+$ indicates the positive part of the expression inside the brackets and subscripts indicate the percentile. Apart from reinsurance, layers arise in the context of derivative payouts and debt tranches in collaterised debt obligations \citep{mandel2012role}. As the interest is in rank dependence, replace the above quantities with ranks and the 95\%-96\% layer reduces to
\begin{equation}\label{eg}
\min\{(u-u_{0.95})^+,u_{0.96}-u_{0.95}\} = \min\{(u-\alpha)^+,\de\alpha\}\ ,
\end{equation}
where $u$ is the percentile rank transform and in this example $\alpha=0.95$ and $\de \alpha=0.01$, and note $u_\alpha=\alpha$.

@@

The final expression in \eref{eg} can be written and approximated by, using the familiar infinitesimal notation,
\begin{equation}\label{ld}
\de \min(u,\alpha) \approx I_\alpha(u)\de \alpha  \cq I_\alpha(u) \equiv \left\{\begin{array}{cc}0\ , & u\le\alpha\ ,\\ 1\ ,& u>\alpha\ .\end{array}\right.
\end{equation}
The left hand side expression in \eref{ld} is called the $\alpha$--layer of $u$. The approximation becomes exact  as $\de\alpha \rightarrow 0$. The $\alpha$--layer of $u$ is an infinitesimally small increment $\de\alpha$ if $u$ exceeds $\alpha$ and is zero otherwise. Hence the $\alpha$--layer captures the movement in $u$ at $\alpha$ and ignores movements elsewhere.  \fref{flayer} illustrates the $\alpha$--layer of $u$.
\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \resizebox{100mm}{!}{\includegraphics{layer.eps}}
    \end{tabular}
    \caption{Illustration of $\alpha$-layer of $u$, written as $I_\alpha(u)\de \alpha$. The $\alpha$-layer of $u$ is only sensitive to movements in $u$ at $\alpha$ and ignores other movements.  }
    \label{flayer}
  \end{center}
\end{figure}

@@

Using the decomposition of $u$ in \eref{decompose},
\begin{equation}\label{spearman}
\cov(v,u)=\cov\left\{v,\int_0^1I_\alpha(u)\de \alpha \right\}= \int_0^1 \cov\{v,I_\alpha(u)\} \de \alpha \; .
\end{equation}
Therefore the covariance $\cov(v,u)$ is decomposed into infinitely many covariances $\cov\{v,I_\alpha(u)\}$ for $0\le\alpha\le 1$. Each covariance measures the dependence between $v$ and the $\alpha$--layer of $u$. In a reinsurance setting, this covariance measures dependence between a particular layer of a loss (in percentile rank terms) and another factor. Alternatively the covariance captures dependence between $v$ and movements in $u$ at $\alpha$. Both interpretations point to $\cov\{v,I_\alpha(u)\}$ as a measure of local dependence between $v$ and $u$. Similar to $\rho$, scaling this covariance with the same when $v=u$ leads to the definition of layer dependence
\begin{equation}\label{ellalpha}
\ell_\alpha \equiv \frac{\cov\{v,I_\alpha(u)\}}{\cov\{u,I_\alpha(u)\}}   \;,
\end{equation}
where the denominator
\begin{equation}\label{covui}
\cov\{u,I_\alpha(u)\}=\E[\{u-\E(u)\}I_\alpha(u)]=\int_\alpha^1 \left(u  - \frac{1}{2}\right)\de \alpha = \frac{\alpha(1-\alpha)}{2}\ .
\end{equation}
Combining \eref{rho}, \eref{spearman}, \eref{ellalpha}  and \eref{covui} yields
\begin{equation}\label{wtdavg}
\rho = 12\int_0^1 \ell_\alpha \cov\{u,I_\alpha(u)\} \de \alpha =  \int_0^1 \ell_\alpha 6\alpha(1-\alpha) \de \alpha \; .
\end{equation}

@@

@@

In \sref{sldcurve} and \sref{sdecompose} it is illustrated and shown  that  $\ell_\alpha=1$ for a particular $\alpha$ occurs if and only if $I_\alpha(u)$ and $I_\alpha(v)$ are both simultaneously 1 or 0: that $u$ and $v$ are both  either greater than $\alpha$ or less than $\alpha$.  Additionally $\ell_\alpha=1$ for $c\le\alpha\le d$ implies $u=v$ over $c\le u,v\le d$. The results weaken as $\ell_\alpha$ decreases from 1.

The following is an alternative expression for $\ell_\alpha$:
\begin{equation}\label{gapexp}
\ell_\alpha = \frac{\E(v|u>\alpha)-\E(v|u\leq \alpha)}{\E(u|u>\alpha)-\E(u|u\leq \alpha)}
=2 \left\{\E(v|u>\alpha)-\E(v|u\leq \alpha)\right\}\;.
\end{equation}
Proofs are in an appendix. The middle expression in \eref{gapexp} is the expected change in $v$ relative to the expected change in $u$ when $u$ crosses $\alpha$. The latter is $0.5$ for all $\alpha$, yielding the final expression in \eref{gapexp}. Hence large $\ell_\alpha$ implies $v$ is sensitive to movements in $u$ across $\alpha$, indicating strong dependence between $v$ and $u$ at $\alpha$. When $\ell_\alpha=0$, $v$ is unchanged on average when $u$ crosses $\alpha$, hence $u$ and $v$ are not dependent at $\alpha$.

@@

\section{Layer dependence, discordance and dispersion}\label{sdecompose}

Again assuming the copula $C$ of $(u,v)$ is exchangeable, then layer dependence $\ell_\alpha$ can be written as
\begin{equation}\label{decomposition}
\ell_\alpha = 1-2(1+\gamma_\alpha) \delta_\alpha \;,
\end{equation}
where
$$
\gamma_\alpha \equiv \cor\{1-I_\alpha(u),I_\alpha(v)\} =\cor\{I_\alpha(u),1-I_\alpha(v)\}=\frac{\alpha^2-C(\alpha,\alpha)}{\alpha(1-\alpha)}\;,
$$
and (** can the next conditioning be written in terms of $I_\alpha(u)$ etc the only way I can think of is $|I_\alpha(u)-I_\alpha(v)|=1$ which is equally messy?**)
$$
\delta_\alpha \equiv\E\left\{(|u-v|)|(u-\alpha)(v-\alpha)<0\right\}  \; .
$$
A proof of \eref{decomposition} is given in the appendix. Equation \eref{decomposition}, as shown below, explains the behaviour of layer dependence curves in \fref{fillustration} where $\ell_\alpha$ for any $\alpha$ increases as the scatter become more tightly concentrated  around the point $(\alpha,\alpha)$, that is lower discordance and dispersion relative to the same point.

@@

\cite{joe1997multivariate} defines coefficients of lower and upper tail dependence as
$$
\lambda_L \equiv \lim_{\alpha\downarrow 0} \p(v\leq \alpha |u\leq \alpha) \cq
\lambda_U \equiv \lim_{\alpha\uparrow 1} \p(v>\alpha |u>\alpha) \;.
$$
Unit coefficients indicate perfect positive tail dependence, and occur if and only if $u$ and $v$ converge simultaneously to 0 (lower tail) or 1 (upper tail). Coefficients of negative tail dependence replace $v\leq \alpha$ and $v>\alpha$ in the above expressions with $v>1-\alpha$ and $v\leq 1-\alpha$, respectively.  \cite{sweeting2013calculating} discusses the drawback of these coefficients and suggests a modification by weakening the limits, yielding links to tail concentration functions discussed below.

(** I'm not completely comfortable with this **) The following shows $\lambda_L=1$ is equivalent to $\ell_0=1$ and $\lambda_U=1$ is equivalent to $\ell_1=1$.   This follows since  $\ell_1=1$ if and only if $\E(v|u=1)=1$. Hence $\ell_1=1$ if and only if $u=1$ implies $v=1$, which is equivalent to $\lambda_U=1$.
From \eref{gapexp}, $\ell_1=2\E(v|u=1)-1$ implying In addition $\ell_0=1$ if and only if $u=0$ implies $v=0$, which is equivalent to $\lambda_L=1$. Similar proofs apply to perfect negative tail dependence.
Hence the measures of  \cite{joe1997multivariate} and layer dependence characterise perfect tail dependence equivalently.

\subsection{Tail concentration function}

Tail concentration \citep{venter2002tails} is a local dependence measure formed from lower and upper conditional tail probabilities. Similar to layer dependence curves, tail concentration functions describe the dependence structure of copulas and are used to identify families of copulas. The following shows an intimate connection between layer dependence and tail concentration. In addition layer dependence refines tail concentration by incorporating additional information on dispersion defined in \sref{sdecompose}.

The tail concentration function at $\alpha$ is:
$$
\tau_\alpha  \equiv \left\{\begin{array}{cc}\p(v\leq \alpha|u\leq \alpha)\ ,  & \alpha\leq 0.5\ \\
\p(v>\alpha|u>\alpha)\ ,& \alpha>0.5 \ \end{array}\right.
=\left\{\begin{array}{cc}\frac{C(\alpha,\alpha)}{\alpha}\ , & \alpha\leq 0.5\ \\
\frac{1-2\alpha+C(\alpha,\alpha)}{1-\alpha}\ ,& \alpha>0.5 \ \end{array}\right. \;.
$$
Higher $\tau_\alpha$ implies $u$ and $v$ are more likely to fall in the same tail -- lower tail if $\alpha\leq 0.5$ and upper tail if $\alpha>0.5$. In addition taking limits $\alpha\rightarrow 0$ and $\alpha\rightarrow 1$ of $\tau_\alpha$ yield coefficients of tail dependence $\lambda_L$ and $\lambda_U$ discussed above. Properties of tail concentration and its applications to distinguish families of copulas are  discussed in \cite{durante2014copulas}.

To show the connection between layer dependence and tail concentration, first standardise $\tau_\alpha$ by subtracting $\tau_\alpha^0$, its value under independence and dividing by the difference between $\tau_\alpha^+$, its value if $u=v$, and $\tau_\alpha^0$:
$$
\tau_\alpha^*=\frac{\tau_\alpha-\tau_\alpha^0}{\tau_\alpha^+-\tau_\alpha^0} = \frac{C(\alpha,\alpha)-\alpha^2}{\alpha(1-\alpha)}
$$
where
$$
\tau_\alpha^0=\left\{\begin{array}{cc}\alpha \ , & \alpha\leq 0.5\ \\
1-\alpha \ ,& \alpha>0.5 \ \end{array}\right.
\cq
\tau_\alpha^+=1 \;.
$$
Then $\tau_\alpha^*$ equals 1 or 0, if $u=v$ or  if $u$ and $v$ are independent, respectively.

Note the final expression for $\tau_\alpha^*=-\gamma_\alpha$ where $\gamma_\alpha$ measures discordance between $u$ and $v$ at $\alpha$ as discussed in \sref{sdecompose}. Combining this result with \eref{decomposition} yields
$$
\ell_\alpha=1-2\delta_\alpha(1-\tau_\alpha^*) = 1-2\delta_\alpha + 2\delta_\alpha \tau_\alpha^*
$$
where $\delta_\alpha$ as defined in \sref{sdecompose} measures average dispersion between discordant $u$ and $v$ at $\alpha$. Hence given $\alpha$, $\ell_\alpha$ is increasing in $\tau_\alpha^*$ and $\tau_\alpha$. It is also straightforward to show that $\ell_\alpha=1$ is equivalent to $\tau_\alpha^*=1$.

Hence there is an intimate connection between tail concentration and layer dependence. However tail concentration (in its standardised form) is only one of two factors forming layer dependence. The other factor is the dispersion between discordant points, measured by $\delta_\alpha$. Using the right panel of \fref{finterpretation}, $\tau_\alpha$ or $\tau_\alpha^*$ only reflects the number of discordant points at $\alpha$, whereas layer dependence combines this information and average dispersion between discordant points. Hence layer dependence refines tail concentration in two ways: first by standardising and then including further information on dispersion.

@@

A direct consequence of the copula integration result \eref{copula} is layer dependence preserves convex combinations of copulas, as follows.

Suppose $(u^*,v^*)$ is bivariate uniform with $\alpha$--layer dependence $\ell_\alpha^*$ and copula $C^*$. Then a mixture of $(u,v)$ and $(u^*,v^*)$ yielding the copula $\pi C+(1-\pi)C^*$ where $0\le \pi\le 1$ has layer dependence $\pi\ell_\alpha+(1-\pi)\ell_\alpha^*$. Hence layer dependence preserves convex combinations of copulas. The proof follows directly from \eref{copula}.

@@

Since $\E(v)=\alpha\E(v|u\leq \alpha)+(1-\alpha)\E(v|u>\alpha)$ it is straightforward to show from \eref{gapexp} that
$$
\ell_\alpha = \frac{\E(v|u> \alpha)-\E(v)}{\E(u|u> \alpha)-\E(u)} = \frac{\E(v|u\leq \alpha)-\E(v)}{\E(u|u\leq \alpha)-\E(u)} \;,
$$
the gap between upper or lower conditional tail expectations of $v$ and the unconditional expectation. Denominators are again scaling factors ensuring $\ell_\alpha=1$ if $u$ and $v$ are comonotonic and $\ell_\alpha=-1$ if countermonotonic.

@@

If the copula of $u$ and $v$ is not exchangeable, $C(u,v)\neq C(v,u)$, then layer dependence changes when $v$ is decomposed into layers rather than $u$:
$$
\frac{\cov\{v,I_\alpha(u)\}}{\cov\{u,I_\alpha(u)\}}
\neq \frac{\cov\{u,I_\alpha(v)\}}{\cov\{v,I_\alpha(v)\}}  \;,
$$
that is, the dependence between $v$ and $\alpha$--layer of $u$ differs from the dependence between $u$ and $\alpha$--layer of $v$.

An analogous property applies to least squares regression: regressing $v$ on $u$ is different from regressing $u$ on $v$ if the joint distribution is not exchangeable.

@@

For various reasons it may still be necessary to refer to measures of overall dependence such as Spearman's correlation. However as discussed in \sref{sintroduction}, Spearman's correlation dismisses tail dependence and may be a less appropriate measure when tail dependence is important such as in insurance and finance. The following formulates alternative measures of overall dependence by taking different weighted averages of layer dependence.

>>>>

Dependence between two random variables generally varies with percentile. For example extreme movements in  stock markets are likely to be highly related  whereas minor fluctuations may be relatively independent.  Catastrophes create significant insurance losses for several classes of business at the same time, while attritional losses are typically weakly dependent.

@@

This paper introduces, illustrates and analyzes an alternate local dependence measure called  ``layer dependence." Layer dependence is the covariance between a random variable and a single ``layer" of another. Layer dependence is also the ``gap" between upper and lower conditional tail expectations. Layer dependence is calculated  from the copula underlying the joint distribution.  Hence of interest is rank dependence rather than dependence between random variables in their original scale: the latter is often distorted by marginal distributions.

@@

Calculating layer dependence at the first instance extracts essential and interpretable information -- the dependence structure. This is often more informative than positing parametric copula, as the implication of its parametric form and parameters on the dependence structure is indirect. Computed layer dependence facilitates the selection and fitting of an appropriate copula.

@@

A familiar construct in the study of bivariate dependence is Spearman's correlation \citep{embrechts2002correlation} defined as the linear correlation between ranks of two random variables. Rank dependence avoids distortion arising from marginal distributions as with for example Pearson's correlation \citep{mcneil2005qrm} measuring the degree of linear relationship between random variables in their original scale. Spearman's correlation can also be applied to estimate copula parameters using the method of moments (\cite{kojadinovic2010comparison}, \cite{bouye2000copulas}). However Spearman's correlation suffers from shortcomings and, as an aggregate measure, is inappropriate for assessing local dependence when dependence varies across the joint distribution including the tails.

Another familiar construct, in reinsurance, is a loss layer \citep{wang1995insurance}.  For example the 95\%--96\% layer of a random continuous loss $x$ is the portion of $x$ between its 95th and 96th percentile
$$
\min\{(x-x_{0.95})^+,x_{0.96}-x_{0.95}\}\ ,
$$
where $()^+$ indicates the positive part of the expression inside the brackets and subscripts indicate the percentile. Layers also arise in the context of derivative payouts and debt tranches in collaterised debt obligations \citep{mandel2012role}.  With rank dependence, the above quantities the 95\%--96\% layer reduces to
\begin{equation}\label{eg}
\min\{(u-u_{0.95})^+,u_{0.96}-u_{0.95}\} = \min\{(u-\alpha)^+,\de\alpha\}\ ,
\end{equation}
where $u$ is the percentile rank  and in this example $\alpha=0.95$ and $\de \alpha=0.01$.  Note $u_\alpha=\alpha$.

@@

The final expression in \eref{eg} can be written and approximated by, using the familiar infinitesimal notation,
\begin{equation}\label{ld}
-\de (u-\alpha)^+ \approx I_\alpha(u)\de \alpha  \cq I_\alpha(u) \equiv \left\{\begin{array}{cc}0\ , & u\le\alpha\ ,\\ 1\ ,& u>\alpha\ .\end{array}\right.
\end{equation}
The left hand side expression in \eref{ld} is called the $\alpha$--layer of $u$. The approximation becomes exact  as $\de\alpha \rightarrow 0$. The $\alpha$--layer of $u$ is an infinitesimally small increment $\de\alpha$ if $u$ exceeds $\alpha$ and is zero otherwise. Hence the $\alpha$--layer captures the movement in $u$ at $\alpha$ and ignores movements elsewhere.  \fref{flayer} illustrates the $\alpha$--layer of $u$.
\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \resizebox{100mm}{!}{\includegraphics{layer.eps}}
    \end{tabular}
    \caption{Illustration of $\alpha$-layer of $u$, written as $I_\alpha(u)\de \alpha$. The $\alpha$-layer of $u$ is only sensitive to movements in $u$ at $\alpha$ and ignores other movements. (*** Maybe u axis should stop at 1 *** }
    \label{flayer}
  \end{center}
\end{figure}

@@

Using the decomposition of $u$ in \eref{decompose},
\begin{equation}\label{spearman}
\cov(v,u)=\cov\left\{v,\int_0^1I_\alpha(u)\de \alpha \right\}= \int_0^1 \cov\{v,I_\alpha(u)\} \de \alpha \; .
\end{equation}
Hence the covariance $\cov(v,u)$ can be thought of as the sum  into infinitely many covariances $\cov\{v,I_\alpha(u)\}$ for $0\le\alpha\le 1$. Each covariance in the sum measures the dependence between $v$ and the $\alpha$--layer of $u$. In a reinsurance setting, this covariance measures dependence between a particular layer of a loss (in percentile rank terms) and another factor. Alternatively the covariance captures dependence between $v$ and movements in $u$ at $\alpha$.

 Similar to $\rho$, scaling the layer covariances  $\cov\{v,I_\alpha(u)\}$ with the same when $v=u$ leads to the definition of layer dependence
\begin{equation}\label{ellalpha}
\ell_\alpha \equiv \frac{\cov\{v,I_\alpha(u)\}}{\cov\{u,I_\alpha(u)\}}   \;,
\end{equation}
where the denominator
\begin{equation}\label{covui}
\cov\{u,I_\alpha(u)\}=\E[\{u-\E(u)\}I_\alpha(u)]=\int_\alpha^1 \left(u  - \frac{1}{2}\right)\de \alpha = \frac{\alpha(1-\alpha)}{2}\ .
\end{equation}
Combining \eref{rho}, \eref{spearman}, \eref{ellalpha}  and \eref{covui} yields
\begin{equation}\label{wtdavg}
\rho = 12\int_0^1 \ell_\alpha \cov\{u,I_\alpha(u)\} \de \alpha =  \int_0^1 \ell_\alpha 6\alpha(1-\alpha) \de \alpha \; .
\end{equation}

@@

In \sref{sldcurve} and \sref{sdecompose} it is illustrated and shown  that  $\ell_\alpha=1$  occurs if and only if $I_\alpha(u)$ and $I_\alpha(v)$ are both simultaneously 1 or 0: that is $u$ and $v$ are both  either greater than $\alpha$ or less than $\alpha$.  Additionally $\ell_\alpha=1$ for $c\le\alpha\le d$ implies $u=v$ over $c\le u,v\le d$.

@@

The following is an alternative expression for $\ell_\alpha$:
\begin{equation}\label{gapexp}
\ell_\alpha = \frac{\E(v|u>\alpha)-\E(v|u\leq \alpha)}{\E(u|u>\alpha)-\E(u|u\leq \alpha)}
=2 \left\{\E(v|u>\alpha)-\E(v|u\leq \alpha)\right\}\;.
\end{equation}
Proofs are in an appendix. The middle expression in \eref{gapexp} is the expected change in $v$ relative to the expected change in $u$ when $u$ crosses $\alpha$. The latter is $0.5$ for all $\alpha$, yielding the final expression in \eref{gapexp}. Hence large $\ell_\alpha$ implies $v$ is sensitive to movements in $u$ across $\alpha$, indicating strong dependence between $v$ and $u$ at $\alpha$. When $\ell_\alpha=0$, $v$ is unchanged on average when $u$ crosses $\alpha$.

@@

\section{Discordance and dispersion}\label{sdecompose}

Layer dependence is intimately connected to the discordance and dispersion.
Assuming the copula $C$ of $(u,v)$ is exchangeable then
\begin{equation}\label{decomposition}
\ell_\alpha = 1-2(1+\gamma_\alpha) \delta_\alpha \;,
\end{equation}
where
$$
\gamma_\alpha \equiv \cor\{1-I_\alpha(u),I_\alpha(v)\} =\cor\{I_\alpha(u),1-I_\alpha(v)\}=\frac{\alpha^2-C(\alpha,\alpha)}{\alpha(1-\alpha)}\;,
$$
and
$$
\delta_\alpha \equiv\E\left\{(|u-v|)|(u-\alpha)(v-\alpha)<0\right\}  \; .
$$
A proof of \eref{decomposition} is  in the appendix. Equation \eref{decomposition}, as shown below, explains the behaviour of layer dependence curves in \fref{fillustration} where $\ell_\alpha$ for any $\alpha$ increases as the scatter become more tightly concentrated  around the point $(\alpha,\alpha)$, that is lower discordance and dispersion relative to the same point.

@@

\cite{joe1997multivariate} defines coefficients of lower and upper tail dependence as
$$
\lambda_L \equiv \lim_{\alpha\downarrow 0} \p(v\leq \alpha |u\leq \alpha) \cq
\lambda_U \equiv \lim_{\alpha\uparrow 1} \p(v>\alpha |u>\alpha) \;.
$$
Unit coefficients indicate perfect positive tail dependence, and occur if and only if $u$ converging to 0 or 1 implies the same for $v$. Coefficients of negative tail dependence replace $v\leq \alpha$ and $v>\alpha$ in the above expressions with $v>1-\alpha$ and $v\leq 1-\alpha$, respectively.  \cite{sweeting2013calculating} discusses the drawback of these coefficients and suggests a modification by weakening the limits, yielding links to tail concentration functions discussed below.

From \eref{gapexp}, $\ell_0=\lim_{\alpha\downarrow 0} \ell_\alpha=1-2\E(v|u=0)$ which equals $1$ if and only if $\E(v|u=0)=0$, or $v$ approaching 0 if the same happens to $u$. Hence $\ell_0=1$ is equivalent to $\lambda_L=1$. Similarly $\ell_1=2\E(v|u=1)-1=1$ if and only if $u=1$ implies $v=1$, which is equivalent to $\lambda_U=1$. Similar results apply to perfect negative tail dependence.

Hence layer dependence characterises perfect tail dependence in the same way as the measures of  \cite{joe1997multivariate}: variables attaining their maximum or minimum values simultaneously.

Tail concentration \citep{venter2002tails} is a local dependence measure formed from lower and upper conditional tail probabilities. Similar to layer dependence, tail concentration functions describe the dependence structure of copulas and are used to identify families of copulas. The following shows the connection between layer dependence and tail concentration and shows layer dependence refines tail concentration by incorporating additional information on dispersion defined in \sref{sdecompose}.

The tail concentration at $\alpha$ is
$$
\tau_\alpha  \equiv \left\{\begin{array}{lr}\p(v\leq \alpha|u\leq \alpha)=\frac{C(\alpha,\alpha)}{\alpha}\ ,  & \alpha\leq 0.5\ , \\
\p(v>\alpha|u>\alpha)=\frac{1-2\alpha+C(\alpha,\alpha)}{1-\alpha}\ ,& \alpha>0.5\ .\end{array}\right.
$$
A larger $\tau_\alpha$ implies $u$ and $v$ are more likely to fall in the same tail -- lower tail if $\alpha\leq 0.5$ and upper tail if $\alpha>0.5$. Letting $\alpha\rightarrow 0$ and $\alpha\rightarrow 1$  yields the coefficients of tail dependence $\lambda_L$ and $\lambda_U$ discussed above. Properties of tail concentration and its applications to distinguish families of copulas are  discussed in \cite{durante2014copulas}.

To show the connection between layer dependence and tail concentration, first standardise $\tau_\alpha$ by subtracting $\tau_\alpha^0$, its value under independence and dividing $\tau_\alpha^+-\tau_\alpha^0$ where $\tau_\alpha^+$ is the value of $\tau_\alpha$ if $u=v$, :
$$
\tau_\alpha^*=\frac{\tau_\alpha-\tau_\alpha^0}{\tau_\alpha^+-\tau_\alpha^0} = \frac{C(\alpha,\alpha)-\alpha^2}{\alpha(1-\alpha)}
\cq
\tau_\alpha^0=\left\{\begin{array}{cc}\alpha \ , & \alpha\leq 0.5\ \\
1-\alpha \ ,& \alpha>0.5 \ \end{array}\right.
\cq
\tau_\alpha^+=1 \;.
$$
Then $\tau_\alpha^*$ equals 1 or 0, if $u=v$ or  if $u$ and $v$ are independent, respectively.

Hence $\tau_\alpha^*=-\gamma_\alpha$ where $\gamma_\alpha$ measures discordance between $u$ and $v$ at $\alpha$ as discussed in \sref{sdecompose}. Combining this result with \eref{decomposition} yields
$$
\ell_\alpha=1-2\delta_\alpha(1-\tau_\alpha^*) = 1-2\delta_\alpha + 2\delta_\alpha \tau_\alpha^*
$$
where $\delta_\alpha$ as defined in \sref{sdecompose} measures average dispersion between discordant $u$ and $v$ at $\alpha$.  Note $\ell_\alpha$ is increasing in $\tau_\alpha^*$ and $\tau_\alpha$. It is also straightforward to show that $\ell_\alpha=1$ is equivalent to $\tau_\alpha^*=1$.

Hence  tail concentration and layer dependence are closely connected. Tail concentration (in its standardised form) is only one of two factors forming layer dependence. The other factor is the dispersion between discordant points, measured by $\delta_\alpha$. Using the right panel of \fref{finterpretation}, $\tau_\alpha$ or $\tau_\alpha^*$ only reflects the number of discordant points at $\alpha$, whereas layer dependence combines this information and average dispersion between discordant points. Hence layer dependence refines tail concentration by standardising and  including  information on dispersion.

@@

Layer dependence is preserved under convex combinations of copulas.  Suppose $(u^*,v^*)$ has  $\alpha$--layer dependence $\ell_\alpha^*$ and copula $C^*$. Then a mixture $\pi(u,v)+(1-\pi)(u^*,v^*)$ has  copula $\pi C+(1-\pi)C^*$ where $0\le \pi\le 1$ and layer dependence $\pi\ell_\alpha+(1-\pi)\ell_\alpha^*$. Hence layer dependence preserves convex combinations of copulas. The proof follows directly from \eref{copula}.

@@

Since $\E(v)=\alpha\E(v|u\leq \alpha)+(1-\alpha)\E(v|u>\alpha)$ it follows from \eref{gapexp} that
$$
\ell_\alpha = \frac{\E(v|u> \alpha)-\E(v)}{\E(u|u> \alpha)-\E(u)} = \frac{\E(v|u\leq \alpha)-\E(v)}{\E(u|u\leq \alpha)-\E(u)} \;,
$$
the gap between upper or lower conditional tail expectations of $v$ and the unconditional expectation. Denominators are again scaling factors ensuring $\ell_\alpha=1$ if $u$ and $v$ are comonotonic and $\ell_\alpha=-1$ if countermonotonic.

@@

If the copula of $u$ and $v$ is not exchangeable, $C(u,v)\neq C(v,u)$, then layer dependence changes when $v$ is decomposed into layers rather than $u$:
$$
\frac{\cov\{v,I_\alpha(u)\}}{\cov\{u,I_\alpha(u)\}}
\neq \frac{\cov\{u,I_\alpha(v)\}}{\cov\{v,I_\alpha(v)\}}  \;,
$$
that is, the dependence between $v$ and $\alpha$--layer of $u$ differs from the dependence between $u$ and $\alpha$--layer of $v$.
This is similar to  regression where the regression of $y$ on $x$ is  not equivalent to the regression of $x$ on $y$ unless the  joint distribution is  exchangeable.

@@

It may still be convenient to use an overall dependence measure such as Spearman's correlation. However as discussed in \sref{sintroduction}, Spearman's correlation downplays tail dependence and may be inappropriate when tail dependence is critical. The following formulates alternative measures of overall dependence by taking different weighted averages of layer dependence:
$$
\rho_W\equiv  \int_0^1 w_\alpha \ell_\alpha \de \alpha
= \int_0^1 w_\alpha \frac{ \cov\{v,I_\alpha(u)\}}{\cov\{u,I_\alpha(u)\}} \de \alpha=\cov\{v,W(u)\} \;,
$$
where
$$
W(u)\equiv \int_0^1 \frac{w_\alpha I_\alpha(u)}{\cov\{u,I_\alpha(u)\}}\de\alpha
= 2\int_0^u \frac{w_\alpha }{\alpha(1-\alpha)}\de\alpha
\cq \int_0^1 w_\alpha \de\alpha=1 \;.
$$
The weighting function $w_\alpha$ indicates the weight or importance of the $\alpha$--layer of $u$.

Hence  $\rho_W$ is the  covariance between $v$ and a transformation of $u$. For Spearman's correlation $\rho$, $w_\alpha=6\alpha(1-\alpha)$ and $W(u)=12u$. Even though the expression for $\rho_W$ is asymmetric in $u$ and $v$, the result is identical when $u$ and $v$ are switched if the copula $C$ is exchangeable.
